---
title: "Image Preprocessing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(readr)
library(fs)
library(stringr)
library(dplyr)
library(colorspace)
library(sf)
library(binaryLogic)
library(sp)
library(velox)
library(readr)
library(glue)
library(dplyr)
```

# Apply UDM

* The Unusable data mask from Planet described [here](https://github.com/planetlabs/notebooks/blob/master/jupyter-notebooks/udm/udm.ipynb)
* Most conservative - only keep pixels where the value of the UDM is 0, i.e. there are no flags. 
* UDMs are saved in `metadata/download_udm/


```{r}
data_dir <- "/nfs/khondula-data/planetmethane"
analytic_sr_dir <- glue("{data_dir}/download_analytic_sr")
myimg_files_nfs <- basename(list.files(analytic_sr_dir))
myimg_ids <- gsub("_analytic_sr.tiff", "", myimg_files_nfs)
udm_dir <- glue("{data_dir}/metadata/download_udm")
all_udms <- basename(list.files(udm_dir))
myudms_ids <- gsub("_udm.tiff", "", all_udms)

message(glue("You have {length(myimg_ids)} images"))
message(glue("You have {length(myudms_ids)} UDMs"))
# myimg_ids[which(!myimg_ids %in% myudms_ids)]
```

Mask the image to set all values as NA that are not 0 in the unusable data mask

First make a vector of the images that are in data/ but not in udm_masked_imgs

```{r ids_to_mask}
masked_imgs_dir <- glue("{data_dir}/udm_masked_imgs")
masked_imgs <- list.files(masked_imgs_dir, pattern = ".grd")
masked_imgs_ids <- tools::file_path_sans_ext(masked_imgs)
message(glue("{length(masked_imgs_ids)} udm masked images"))

imgIDs_to_mask <- myimg_ids[!myimg_ids %in% masked_imgs_ids]
message(glue("{length(imgIDs_to_mask)} udm masked images"))
message("Not masked: ", paste(imgIDs_to_mask, collapse = ", "))

imgIDs_to_mask_w_udms <- imgIDs_to_mask[imgIDs_to_mask %in% myudms_ids]
imgIDs_to_mask_w_udms
```

Apply UDM mask to vector you just created - IDs with UDMs that arent in udm_masked_imgs folder:

```{r udm_func_defn}
# imgID <- imgIDs_to_mask_w_udms[1]

apply_udm_mask <- function(imgID){
  
  data_dir <- "/nfs/khondula-data/planetmethane"
  analytic_sr_dir <- glue("{data_dir}/download_analytic_sr")
  myimg_files_nfs <- list.files(analytic_sr_dir, full.names = TRUE)
  myimg_file <- grep(imgID, myimg_files_nfs, value = TRUE)
  myimg <- raster::stack(myimg_file)

  udm_dir <- glue("{data_dir}/metadata/download_udm")
  all_udms <- list.files(udm_dir, full.names = TRUE)
  myudms_ids <- gsub("_udm.tiff", "", all_udms)
  myudm_file <- grep(imgID, all_udms, value = TRUE)
  myimg_udm <- raster(myudm_file)

  masked_filename = glue("/nfs/khondula-data/planetmethane/udm_masked_imgs/{imgID}.grd")
  masked_stack <- mask(myimg, 
                       myimg_udm, 
                       inverse = TRUE, 
                       maskvalue = 0, 
                       updatevalue = NA,
                     filename = masked_filename, 
                     overwrite = TRUE)  
}

```

Run function locally 

```{r}
# apply_udm_mask(imgID = imgIDs_to_mask_w_udms[1])
# purrr::walk(imgIDs_to_mask_w_udms[], ~apply_udm_mask(.x))
```

# Add Bands

First figure out which images will be used for analysis and need the additional bands added

```{r}
masked_imgfiles <- fs::dir_ls(glue("{data_dir}/udm_masked_imgs/"), glob = "*.grd")
masked_ids <- gsub(".grd", "", basename(masked_imgfiles))
message(paste(length(masked_ids), "udm masked images"))

addbands_files <- fs::dir_ls(glue("{data_dir}/imgs_for_analysis"))
addbands_ids <- gsub(".tif", "", basename(addbands_files))
message(paste(length(addbands_ids), "images for analysis with bands added"))

# 
needbands_files <- masked_imgfiles[!masked_ids %in% addbands_ids]

message(paste(length(needbands_files), "udm masked images without bands added"))

```


```{r addbands_func_defn}
masked_imgfile <- needbands_files[1]
# masked_imgfile <- masked_imgfiles[1]

save_allbands <- function(masked_imgfile){
  # read in original 4 bands
    data_dir <- "/nfs/khondula-data/planetmethane"
  r_stack <- raster::stack(masked_imgfile)
  # get imgID
  imgID <- basename(masked_imgfile) %>% gsub(".grd", "", .)
    
  new_filename <- glue("{data_dir}/imgs_for_analysis/{imgID}.tiff")
  # add NDVI layer
  r_stack$NDVI <- raster::overlay(r_stack[[3]], r_stack[[4]],
                                   fun = function(b3, b4){return((b4-b3)/(b4+b3))})
                                   # filename = new_filename)
  # add NDWI layer
  r_stack$NDWI <- raster::overlay(r_stack[[2]], r_stack[[4]],
                                   fun = function(b2, b4){return((b2-b4)/(b2+b4))})
                                   # filename = new_filename)
  # get vectors of band values
  b1_blue <- velox::velox(r_stack)$as.matrix(band = 1) %>% as.vector()
  b2_green <- velox::velox(r_stack)$as.matrix(band = 2) %>% as.vector()
  b3_red <- velox::velox(r_stack)$as.matrix(band = 3) %>% as.vector()
  b4_NIR <- velox::velox(r_stack)$as.matrix(band = 4) %>% as.vector()
  
  # convert to data frame
  rgbvals <- data.frame("b1_blue" = b1_blue, 
                        "b2_green" = b2_green, 
                        "b3_red" = b3_red) %>%
    dplyr::mutate(cellid = 1:ncell(r_stack)) 
  
  rgbvals_narm <- rgbvals %>% 
    dplyr::filter(!is.na(b1_blue), !is.na(b2_green), !is.na(b3_red))
  # calculate hue saturation value
  hsvvals <- rgb2hsv(r = rgbvals_narm[,"b3_red"],
                   g = rgbvals_narm[,"b2_green"], 
                   b = rgbvals_narm[,"b1_blue"], 
                   maxColorValue = max(rgbvals_narm[,1:3]))
  # calculate chroma luminance hue2
  labvals <- colorspace::RGB(rgbvals_narm[,"b3_red"], 
                           rgbvals_narm[,"b2_green"],
                           rgbvals_narm[,"b1_blue"]) %>% as("polarLAB")
  
  # get all values together in a dataframe
  hsv_df <- hsvvals %>% t() %>% as.data.frame() %>% 
    dplyr::mutate(cellid = rgbvals_narm$cellid)

  labvals_df <- labvals@coords %>% as.data.frame() %>% 
    dplyr::mutate(cellid = rgbvals_narm$cellid)
  
  colmetrics_df <- rgbvals %>% 
    dplyr::left_join(hsv_df) %>% 
    dplyr::left_join(labvals_df)

  # define new function to make raster layer from column
  column_to_rasterlayer <- function(colorcol){
    lum_mat <- matrix(colmetrics_df[[colorcol]], nrow(r_stack), ncol(r_stack))
    lum_vx <- velox::velox(lum_mat, extent = extent(r_stack), res = res(r_stack))
    return(lum_vx$as.RasterLayer())}

  r_stack$HUE <- column_to_rasterlayer("h")
  r_stack$SAT <- column_to_rasterlayer("s")
  r_stack$VAL <- column_to_rasterlayer("v")
  r_stack$LUM <- column_to_rasterlayer("L")
  r_stack$CHROMA <- column_to_rasterlayer("C")
  r_stack$HUE2 <- column_to_rasterlayer("H")

  raster::writeRaster(r_stack, new_filename, overwrite = TRUE)    

  }
```

```{r}
# save_allbands(masked_imgfile = masked_imgfiles[3])
# purrr::walk(needbands_files, ~save_allbands(masked_imgfile = .x))
```

